---
title: "Final Project"
author: "Leslie Alhakim"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1: Data Wrangling

```{r data-wrangling}
# Load necessary libraries
library(dplyr)
library(tidyverse)
library(ggplot2)
library(paletteer) 
library(readr)

# Load our dataset and explore the structure of the data
movies <- read_csv("movies.csv")
names(movies)
head(movies) #let's see what the data looks like 

# Cleaning time!
df<- movies %>% mutate(month=word(released,1)) #take month info from released col
df <- df %>% mutate(revenue = gross - budget) #calculate $$ revenue
df <- df %>% select(-released, -runtime) #remove unnecessary cols
names(df)
head(df)
```

```{r save dataframe}
# Save clean data
write_csv(df, "CLEAN.csv")
```

## 2: Visualization Creation

```{r ggplot2}
# PLOT 1: Density Plot
grouped <- df %>%
  mutate(genre = fct_lump_n(as.factor(genre), n = 6)) #little did I know lump would be my new best friend
#group 6 major movies using tidyverse

p1 <- ggplot(data = grouped, aes(x = year)) +
  geom_density(aes(colour = genre), alpha = 0.3, linewidth = .8) +
  labs(
    colour = "Genre",
    x = "Year",
    y = "Density",
    title = "Prevalence in Major Movie Genres From 1980 to 2020"
  ) +
  theme_minimal() +
  scale_colour_paletteer_d("MetBrewer::Austria")

p1
```
This is a density plot of the six major movie genres from 1980 to 2020. I chose this plot because
I thought it would be interesting to see the rise/fall in popularity of certain movie genres.
While this plot would be interesting if it contained all genres in the database, including
19 genres made the density plot seem crowded and difficult to read. Thus, it made sense to 
only include the top 6 most common genres and group the rest into "Other". I used tidyverse 
to lump the six most common genres.
Lastly, I used colour instead of fill to make the plot legible and included all necessary 
headings. Overall, I think this plot gives a good visual of genre trends over the years.

```{r ggplot2}
# PLOT 2: Heat Map
genre_count <- df %>%
  mutate(
    month = factor(month, levels = month.name),
    genre = fct_lump_n(as.factor(genre), n = 10)
  ) %>%
  count(genre, month) #the original plot looked pretty ugly so we need to focus on top 10
#also we need count of movies per genre released each month

p2 <- ggplot(data = genre_count, aes(x = month, y = genre)) +
  geom_tile(aes(fill = n)) +
  scale_fill_viridis_c(name = "Number of Movies") +
  labs(
    x = "Month",
    y = "Genre",
    title = "Movie Releases by Month and Genre From 1980 to 2020"
  ) +
  theme_minimal() +
 theme(
  axis.text.x = element_text(angle = 45, hjust = 1)) #tilt text for better look 

p2
```
This is a heat map of movie releases by genre and month. My reasoning for creating this plot
was to look at movie release trends within different genres by month. For instance,
are more Horror movies released around Halloween? Or maybe more romance movies are released
around Valentine's Day? For this plot, I had to take the count of movies in each genre 
counted by month and then take the top 10 (too many genres overall). Then creating
the scale fill took me a lot of time, but I was able to figure it out. Even though this plot was a bit 
challenging to make (lots of errors), I still don't love the end result. 
I don't love the color palette and the trends aren't as clear as I thought they would be, 
but I think it's still overall a good visual. I did end up rotating the months on the x axis for
readability. 

```{r ggplot2}
# PLOT 3: Box Plot
grouped2 <- df %>%
  mutate(genre = fct_lump_n(as.factor(genre), n = 10)) # top 10 genres

p3 <- ggplot(data = grouped2, aes(x = genre, y = score)) +
  geom_boxplot(aes(fill = genre)) +
  labs(
    fill = "Genre",
    x = "Major Movie Genres",
    y = "Audience Score",
    title = "Audience Ratings by Genre"
  ) +
  theme_minimal() + 
   theme(
  axis.text.x = element_text(angle = 45, hjust = 1)) #tilt text for better look 

p3
```
This is a box plot of major movie genres and information about the ranges of their audience
ratings. I thought it would be interesting to learn more about how each genre is typically rated, 
and what the range of those ratings look like. As a movie fan myself, I've heard that certain
genres typically perform better than others and I made this box plot to test that theory myself.
Making the box plot was not difficult, as we have done this in other assignments. I did only use
the top 10 genres since there are a lot of different genres in this data set. I also tilted the 
x axis text again. Overall, this plot did show me the information I was curious about. 

```{r ggplot2}
# PLOT 4: Bar Chart
top_companies <- df %>%
  group_by(company) %>%
  summarise(total_rev = sum(revenue, na.rm = TRUE)) %>%
  arrange(desc(total_rev)) %>%
  slice_head(n = 10) #find out who's the top ten companies by revenue $$

p4 <- ggplot(data = top_companies, aes(x = reorder(company, total_rev), #reorder
                                       y = total_rev)) +
  geom_col(aes(fill = company)) +
  labs(
    fill = "Company",
    x = "Company Name",
    y = "Total Revenue (Gross - Budget)",
    title = "Top 10 Production Companies by Revenue"
  ) +
  theme_minimal() +
  scale_fill_paletteer_d("rcartocolor::Antique") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
p4
```
This is a bar chart of the top 10 production companies based off revenue made from 1980
to 2020. I chose this visual because I thought it would be interesting to see different
industry giants and who's dominating. I decided to first group the top ten companies
by revenue and then create a ggplot (which we've done before) to display the relation.
I color-coded each company and tilted the text to make the visual readible. I also 
ordered the bar chart for quick analysis on who's making the most/least revenue. 

```{r ggplot2}
# PLOT 5: Line Chart
company_rev <- df %>%
  filter(!is.na(revenue)) %>%
  group_by(company,year) %>%
  summarise(total_rev = sum(revenue), .groups = "drop") 
#so here we need the summary of each company's revenue
#originally, I didn't realize I needed to filter this but without it, it 
#just plots each individual point of revenue for each movie

top_companies <- company_rev %>%
  group_by(company) %>%
  summarise(total = sum(total_rev)) %>%
  slice_max(total, n = 5) %>%
  pull(company)
#and then using that info, we take the top 5 companies (like before) because there are way too many companies

company_rev <- company_rev %>%
  filter(company %in% top_companies) #then we add the top companies into that filtered revenue

p5 <- ggplot(data = company_rev, aes(x = year, y = total_rev)) +
  geom_line(aes(colour = company), alpha = 1, linewidth = .8) +
  labs(
    colour = "Production Company",
    x = "Year",
    y = "Revenue",
    title = "Top 5 Production Companies' Revenue From 1980 to 2020"
  ) +
  theme_gray() + #dark theme for lighter colors
  scale_colour_viridis_d(option = "turbo")

p5 <- p5 + scale_x_continuous(breaks = seq(1980, 2020, by = 3))
#when I showed this plot to a friend, they send it was hard to read so 
#I think expanding the x axis will look better (3 years instead of 5)

p5 
```
This is a line chart of the revenue of the top 5 production companies' revenue over the years.
With the last plot, I got to learn about which companies were making the most revenue, but here
I wanted to learn more about the lifespan of these companies. For instance, was Universal
Pictures always making the most? Which companies were dominating and went down over the years? 
I was especially interested in this information due to the growth of Netflix and how streaming
platforms have changed the game. I did not, however, expect this plot to take this much effort. 
I had to group the revenue made by each company and then again create the top 5 companies based off revenue. 
Then, I used a dark theme so that all colors can be clearly seen. I also changed the x axis to display
every 3 years since the lines were so condensed. I am very satisfied with this end result
and happy I had a second pair of eyes look at this for aesthetics. 

```{r ggplot2}
# PLOT 6: Faceted Scatter Plot
facet_df <- df %>% #save into new df (just in case)
  mutate(
    decade = paste0(floor(year / 10) * 10, "s"), #decades are every 10 years
    #I know 2020 only has one year but listen, it is what it is
    genre = fct_lump_n(as.factor(genre), n = 10)) %>%
  filter(!is.na(revenue), !is.na(year)) #coloring by genre so take top genres based off revenue
#same as we did for companies
    
p6 <- ggplot(data = facet_df, aes(x = year, y = revenue)) +
  geom_point(aes(colour = genre,
                 text = paste(
                   "Movie:", name,
                   "<br>Genre:", genre,
                   "<br>Revenue:", scales::dollar(revenue)),#for later- trust the process
                 alpha=.4)) + #alpha lower to see overlapping points
  facet_wrap(~ decade, scales = "free_x") + #we want each x axis to show ONLY years within decade
  scale_x_continuous(breaks = seq(1980, 2020, by = 2)) + #show by 2 (similar as last vis) 
  labs(
    colour = "Genre",
    x = "Year",
    y = "Revenue",
    title = "Revenue of Different Movies in Each Decade"
  ) +
  theme_minimal() +
  scale_colour_paletteer_d("rcartocolor::Antique")

p6
```
```{r ggplot2-interactive}
# EXTRA: Interactive Plotly of Plot 6
library(plotly)
p7 <- ggplotly(p6, tooltip = "text") #this is what text was for

p7
```
Okay, so I know that creating an Interactive Plotly was not required for the final, BUT I really wanted to 
create a faceted scatter plot of different movies' revenue in each decade to see what were the top
grossing movies in each decade. Using a regular faceted scatter plot, I could not make the colors
be the different movies (too many options), so I needed a tool tip. This is my major visual and what I 
was most interested in, so I think it was worth the extra step. 
For this masterpiece, I had to create facets of each decade, which I did by taking every 10 years
and adding an "s". Then, I took the top 10 genres and grouped them by who made the most revenue. This was
pretty straight forward since I've been using lump throughout this project. I then created the faceted scatter
plot, which we have done for other assignments. 
In preparation of the plotly, I created my tooltip ahead of time with movie, genre, and revenue information.
Then I just converted to plotly and defined my tooltip. I am VERY happy with this last visual because
I think it ties in so many different things I learned this semester. Plus, it's interesting!